% !TEX root = ../thesis.tex

\prefacesection{Acknowledgments}

Without the support of a large community of excellent people, I would not be where I am today.

I would like to thank Professor Steve Rock for his guidance and support over the past years. Professor Rock acted like a finely-tuned regulator for all his students, keeping us on course through all the highs and lows of the PhD. I suppose this isn't surprising, given his background in control theory. I will also always appreciate the freedom he gave us to find our own inspiration and solutions, and for the congenial environment he fostered in the Aerospace Robotics Lab. I couldn't have asked for a better home during my time at Stanford.

I would also like to thank my dissertation committee members. Professor Chris Gerdes gave me my first introduction to control theory back in E 105, and I had the good fortune to assist him with the class in the years that followed. Likewise, Professor Paul Mitiguy tore down everything I thought I knew about dynamics and rebuilt me into a person who dreams in vector calculus. Professor Juan Alonso guided me through my masters program and has always supported my flying robot habit. Special thanks also to Dr. Phyllis Gardner, for serving as committee chair, and for her and her husband Andy's tireless support of my fiancée and me.

A few other faculty members were not on my committee, but were no less influential in my success at Yale and Stanford. Without Professor Ed Carryer's mechatronics classes and continuing advice throughout the years, my desk would be much less cluttered with microcontrollers and wires. I would also be a much less happy and less capable engineer. John Morrell's dedication to his mechanical engineering students at Yale was without parallel, and I will be forever grateful for his mentorship. Dr Jonathan Holloway (and family) made Yale and Calhoun College feel like home away from home, and will always hold a special place in my heart.

The Aerospace Robotics Lab was home at Stanford, and its members were family. Whether bouncing ideas off of each other, solving crossword puzzles, playing scotch toss, or mass lunch migrations, I couldn't have wished for a better group of people to hang out with day in and day out. When I first arrived on campus, the already-established lab members gave me something to aspire to. Peter Kimball, Stephen Russell, Sean Augenstein, Debbie Meduna, Eleanor Crane, Kiran Murthy, Dan Sheinfeld, Andrew Smith: to the person, wickedly smart, and just plain cool individuals. Roland Burton's sharp insights and generosity never cease to amaze. Steve Krukowski can always be counted on to be the life of the lab party. Sarah Houts, Shandor Dektor, and Jose Padial, between traipsing around Europe or Canada, and paddling leaky rubber dinghies, there are too many good memories to count. Ozhan Turgut is possibly the most likable human being ever created, and introduced me to the phenomenon of Turkish Oil Wrestling. Need I say more? And finally, to the new guard, Ashley Clark, David Stonestrom, Aditya Mahajan, I look forward to seeing where you carry the torch.

The engineers and scientists at Monterey Bay Aquarium Research Institute have been an enormous help with providing data for my research, as well as in sharing their technical expertise. Particular thanks to Mike Risi, Brett Hobson, and Dave Caress, as well as the crew of the Western Flyer and all the ROV pilots. 

Thanks to the ARL and Stanford Aero/Astro staff for keeping the ship afloat. It was always a pleasure to have Godwin Zhang in lab. He is institutional wisdom personified. Sherann Ellsworth and Dana Parga were a dream to work with on the administrative side, and tireless advocates for us grad students. Patrick Ferguson, Jay Subramanian and the rest of the front office staff were always there to help navigate the bureaucratic academic waters.

I'd like to thank the rest of my cohort, who for some unfathomable reason decided to study something other than robotic mapping, navigation, and control. Kyle Washabaugh, it's been over a decade. Here's to several more. Zach Vane, Tom Economon, Sean Copeland, George Anderson, Zouhair Mahboubi, I'll always look back on frisbee golf and all the other memories fondly. Except studying for quals. That was miserable. I'm still trying to forget it.

A huge thank you has to go out to Los Salseros de Stanford. From my first week on campus, you have been family. If I were to say everything I wanted to say about each and every one of you, it would  probably double the length of this thesis, so I won't, but you know who you are, and you mean the world to me.

Of course, I have my family to thank for laying a solid foundation on which to build. My parents, John and Ruth Anne, always placed my sister's and my wellbeing above all else, and fostered our curiosity and integrity. For that, I am eternally grateful. My younger sister Ivy's compassion, drive, and dedication to helping others continuously inspires. Finally, thanks to my fiancée Lauryann for putting up with the late nights and lost weekends that come with research. She provides much-needed emotional balance and happiness to my otherwise technical existence, and I am thrilled to dance through life with her. I love you all.
